---
title: "Interactive AI Job Share Map"
author: "Jacob Khaykin"
format:
  html:
    theme: cosmo
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

<style>
/* Page styling */
body {
  background-color: #f7f9fc;
  font-family: 'Lato', sans-serif;
  color: #2c3e50;
  margin: 0;
  padding: 1rem;
}

h1, h2 {
  text-align: center;
  margin-bottom: 1rem;
}

.intro {
  max-width: 800px;
  margin: 0 auto 2rem;
  padding: 1rem;
  background: #ffffff;
  border-radius: 8px;
  box-shadow: 0 2px 6px rgba(0,0,0,0.1);
}

.leaflet-container {
  border: 2px solid #ddd;
  border-radius: 8px;
  margin-bottom: 2rem;
}
</style>

<div class="intro">
Have you ever wondered which U.S. counties lead in AI‑related job postings? 
This interactive map highlights both the average AI share over 2018‑2023 and the percentage‑point change since 2017‑2018.  

- **Average AI Share** shows where AI roles are most concentrated today.  
- **Change in AI Share** reveals the fastest‑growing hotspots of remote‑friendly AI jobs.  

Toggle between layers to explore regional patterns and see how traditional tech hubs compare with emerging counties.
</div>

```{r setup}
library(tidyverse)
library(sf)
library(leaflet)
library(scales)
library(tigris)
``` 

```{r load-data}
# Load and prepare AI data
df <- read_csv("data.csv") %>%
  filter(Year %in% c(2017,2018,2022,2023)) %>%
  mutate(
    ai_intensity = ai / nads,
    period = ifelse(Year %in% c(2017,2018), "early", "late")
  ) %>%
  group_by(COUNTY_FIPS, period) %>%
  summarise(ai_intensity = mean(ai_intensity, na.rm = TRUE), .groups = 'drop') %>%
  pivot_wider(names_from = period, values_from = ai_intensity) %>%
  mutate(
    ai_change_pct = (late - early) * 100,
    avg_ai_pct    = ((early + late)/2) * 100
  )
# Format FIPS codes
df$COUNTY_FIPS <- str_pad(as.character(df$COUNTY_FIPS), 5, pad = '0')
``` 

```{r load-shapefile}
# Quietly load county geometries
options(tigris_use_cache = TRUE)
counties <- suppressMessages(
  tigris::counties(cb=TRUE, year=2020, class='sf')
) %>%
  mutate(COUNTY_FIPS = GEOID) %>%
  st_transform(crs=4326)

# Merge with AI data
map_data <- left_join(counties, df, by='COUNTY_FIPS')
``` 

```{r create-palettes}
# Define bins and colors
bins_avg <- c(0, 0.06, 0.14, 0.23, 0.37, 0.71, Inf)
cols_avg <- c('#ffffcc','#ffeda0','#fed976','#fd8d3c','#e31a1c','#800026')

bins_chg <- c(-5.56, -0.12, 0.00, 0.09, 0.24, 0.57, Inf)
cols_chg <- c('#ffffcc','#ffeda0','#fed976','#fd8d3c','#fc4e2a','#800026')

pal_avg <- colorBin(palette=cols_avg, domain=map_data$avg_ai_pct, bins=bins_avg, na.color='gray90', right=FALSE)
pal_chg <- colorBin(palette=cols_chg, domain=map_data$ai_change_pct, bins=bins_chg, na.color='gray90', right=FALSE)
``` 

```{r render-map}
leaflet(map_data, options=leafletOptions(minZoom=4)) %>%
  setView(lng=-98.5795, lat=39.8283, zoom=4) %>%
  addTiles() %>%
  addPolygons(
    fillColor=~pal_avg(avg_ai_pct), weight=0.3, color='white', fillOpacity=0.75,
    label=~paste0(NAME, ': ', round(avg_ai_pct,1), '%'),
    group='Average AI Share'
  ) %>%
  addLegend(pal=pal_avg, values=~avg_ai_pct, title='Avg AI Share (%)', position='topright', labFormat=labelFormat(suffix='%')) %>%
  addPolygons(
    fillColor=~pal_chg(ai_change_pct), weight=0.3, color='white', fillOpacity=0.75,
    label=~paste0(NAME, ': ', round(ai_change_pct,2), ' pp'),
    group='Change in AI Share'
  ) %>%
  addLegend(pal=pal_chg, values=~ai_change_pct, title='Change in AI Share (pp)', position='bottomright', labFormat=labelFormat(suffix=' pp')) %>%
  addLayersControl(
    overlayGroups=c('Average AI Share', 'Change in AI Share'),
    options=layersControlOptions(collapsed=FALSE)
  )
```
<div class="conclusion" style="max-width:800px;margin:2rem auto;padding:1rem;background:#fff;border-radius:8px;box-shadow:0 2px 6px rgba(0,0,0,0.1);"> **Map Takeaways:** Even outside Silicon Valley, some unexpected corners of America—like Slope County (ND) and Maries County (MO)—have AI roles making up over 5–10% of all job ads. Meanwhile, many rural and industrial counties remain below 1%. Between 2018 and 2023, suburban and remote‑friendly regions saw the fastest growth, underscoring how AI work is decoupling from traditional tech hubs. This uneven landscape highlights both the promise of remote work and the risk that smaller communities could be left behind in the AI economy. </div>